Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations

### 2024-06-06 index.Rmd
add last changes in html

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
```

```{r message=FALSE, warning=FALSE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # whitout S-R
dir4<-here("s4") # 
dir5<-here("s5") # whitout S-R
dir6<-here("s6") # whitout S-R
dir7<-here("s7") # whitout S-R
dirtest<-here("test")
Figs <- here("Figs")# S2 set parametres EMM-2024/23 (Mardones)
```

@@

 SRpar<-ctl1$SR_parms[1:5,c(1:3,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:3,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:3,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))

wb <- createWorkbook()

@@

```{r}
parInit %>%
  kbl(booktabs = T,
      format = "latex",
      position="ht!",
    caption = "\\label{Tab1}Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9)%>% 
  pack_rows(index = c("Mortalidad natural" = 1,
                        "Crecimiento"= 5,
                        "Relación longitud-peso" = 2,
                        "Ojiva de madurez"=2,
                        "Relación stock-recluta"=5,
                        "Capturabilidad"=2,
                        "Selectividad"=4))

@@

```{r eval=FALSE, message=F, include=FALSE}
# Lista de directorios para correro tordos juntos
directorios <- c("s01", 
                 "s1", 
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar

@@

```{r}
total01 <- as.data.frame(base.model01$timeseries[1:50,c(2,5)]) %>%
  mutate(Serie = "No Spatial")
total1 <- as.data.frame(base.model1$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial implicit")
total2 <- as.data.frame(base.model2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ without S-R relation")
total3 <- as.data.frame(base.model3$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation weak (0.9 steepness)")
total4 <- as.data.frame(base.model4$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation strong (0.6 steepness)")
total5 <- as.data.frame(base.model5$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation mid-strong estimated ")
total6 <- as.data.frame(base.model6$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ Ricker S-R relation estimated")
total7 <- as.data.frame(base.model7$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ New set parameters Linf, k and M")

@@

```{r fig.height=7}
SSplotData(base.model01, subplot = 1, 
           fleetnames = c("Fishery", "Survey", "Predator"),
           fleetcol = c("blue","green", "red"))

@@

```{r}
SSplotSelex(base.model01,
            subplots = 1)
SSplotSelex(base.model1,
            subplots = 1)
SSplotSelex(base.model2,
            subplots = 1)
SSplotSelex(base.model3,
            subplots = 1)
SSplotSelex(base.model4,
            subplots = 1)
SSplotSelex(base.model5,
            subplots = 1)
SSplotSelex(base.model6,
            subplots = 1)
SSplotSelex(base.model7,
            subplots = 1)
```

```{r eval=FALSE}
SSplotComps(base.model01,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)

SSplotComps(base.model1,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model2,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model3,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model4,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model5,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model6,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model7,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```

@@

## Residual consistency 
```{r}
SSplotRunstest(base.model01,
               subplots = "len",
               add=T)
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
SSplotRunstest(base.model2,
               subplots = "len",
               add=T)
SSplotRunstest(base.model3,
               subplots = "len",
               add=T)
SSplotRunstest(base.model4,
               subplots = "len",
               add=T)
SSplotRunstest(base.model5,
               subplots = "len",
               add=T)
SSplotRunstest(base.model6,
               subplots = "len",
               add=T)
SSplotRunstest(base.model7,
               subplots = "len",
               add=T)
```

```{r}
SSplotRunstest(base.model01,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model3,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model4,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model5,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model6,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model7,
               subplots = "cpue",
               add=T)
```

@@

```{r}
hci = SSplotHCxval(retroSummary01, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary1, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary2, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary3, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7) 
hci = SSplotHCxval(retroSummary4, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary5, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary6, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary7, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```

@@

```{r}
mvln = SSdeltaMVLN(base.model01,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

@@

## Verosimilitud

@@

## SSplotComparisons()

@@

addWorksheet(wb, "variable_s2")
writeData(wb, "variable_s2", outps1)

# Guardar el workbook
saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
```

@@

comp01_1_7 <- SSplotComparisons(summaryoutput01_1_7,
                  legendlabels = c(
                                   "No Spatial",

@@

```{r}

@@

```{r}
recs01 <- base.model01$recruit$pred_recr
recs1 <- base.model1$recruit$pred_recr
recs2 <- base.model2$recruit$pred_recr
recs3 <- base.model3$recruit$pred_recr
recs4 <- base.model4$recruit$pred_recr
recs5 <- base.model5$recruit$pred_recr
recs6 <- base.model6$recruit$pred_recr
recs7 <- base.model7$recruit$pred_recr

@@

```{r}
dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S01")
dev1 <- base.model1$recruit[1:43,c(1,7)]%>% 
  mutate(Serie = "S1")
dev2 <- base.model2$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S2")
dev3 <- base.model3$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S3")
dev4 <- base.model4$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S4")
dev5 <- base.model5$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S5")
dev6 <- base.model6$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S6")
dev7 <- base.model7$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S7")

>>>>

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 3,
                      fig.width = 5,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
library(readxl)
```

```{r message=FALSE, warning=FALSE, echo=TRUE}
dir01 <- here("s01") # agreggate data (no spatial diferences)
dir1<-here("s1") # Data strata flishery
dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
dir3<-here("s3") # whitout S-R
dir4<-here("s4") # 
dir5<-here("s5") # 
dir6<-here("s6") # 
dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
dirtest<-here("test")
Figs <- here("Figs")# S
```

@@

 SRpar<-ctl1$SR_parms[1:5,c(1:3,7)]
 Qpar<-ctl1$Q_parms[1:2,c(1:3,7)]
 Selpar<-ctl1$size_selex_parms[1:22,c(1:3,7)]
 parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
```

```{r eval=FALSE}
wb <- createWorkbook()

@@

```{r}
parInit %>%
  kbl(booktabs = T,
      format = "html",
      position="ht!",
    caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", 
              full_width = F)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9)%>% 
  pack_rows(index = c("Mortalidad natural" = 1,
                        "Crecimiento"= 5,
                        "Relación longitud-peso" = 2,
                        "Ojiva de madurez"=2,
                        "Relación stock-recluta"=5,
                        "Capturabilidad"=2,
                        "Selectividad"=4))

@@

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
# Lista de directorios para correro tordos juntos
directorios <- c("s01", 
                 "s1", 
                 "s2",
                 "s3",
                 "s4",
                 "s5",
                 "s6",
                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar

@@

```{r eval=FALSE}
total01 <- as.data.frame(base.model01$timeseries[1:50,c(2,5)]) %>%
  mutate(Serie = "No Spatial")
total1 <- as.data.frame(base.model1$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial implicit")
total2 <- as.data.frame(base.model2$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ without S-R relation")
total3 <- as.data.frame(base.model3$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation weak (0.9 steepness)")
total4 <- as.data.frame(base.model4$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation strong (0.6 steepness)")
total5 <- as.data.frame(base.model5$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ S-R relation mid-strong estimated ")
total6 <- as.data.frame(base.model6$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ Ricker S-R relation estimated")
total7 <- as.data.frame(base.model7$timeseries[,c(2,5)]) %>% 
  mutate(Serie = "Spatial w/ New set parameters Linf, k and M")

@@

```{r }
SSplotData(base.model01, subplot = 1, 
           fleetnames = c("Fishery", "Survey", "Predator"),
           fleetcol = c("blue","green", "red"))

@@

```{r fig.height=6}
SSplotSelex(base.model01,
            subplots = 1)
SSplotSelex(base.model1,
            subplots = 1)
SSplotSelex(base.model2,
            subplots = 1)
SSplotSelex(base.model3,
            subplots = 1)
SSplotSelex(base.model4,
            subplots = 1)
SSplotSelex(base.model5,
            subplots = 1)
SSplotSelex(base.model6,
            subplots = 1)
SSplotSelex(base.model7,
            subplots = 1)
```
```{r}
SSplotComps(base.model1,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```

```{r eval=FALSE}
SSplotComps(base.model01,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model1,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model2,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model3,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model4,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model5,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model6,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
SSplotComps(base.model7,
            subplots = 1,
            maxrows = 5,
            maxcols = 5)
```

@@

## Residual consistency

```{r}
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
```

```{r eval=FALSE, fig.width=4, fig.height=5}
SSplotRunstest(base.model01,
               subplots = "len",
               add=T)
SSplotRunstest(base.model1,
               subplots = "len",
               add=T)
SSplotRunstest(base.model2,
               subplots = "len",
               add=T)
SSplotRunstest(base.model3,
               subplots = "len",
               add=T)
SSplotRunstest(base.model4,
               subplots = "len",
               add=T)
SSplotRunstest(base.model5,
               subplots = "len",
               add=T)
SSplotRunstest(base.model6,
               subplots = "len",
               add=T)
SSplotRunstest(base.model7,
               subplots = "len",
               add=T)
```

```{r}
SSplotRunstest(base.model1,
               subplots = "cpue",
               add=T)
```

```{r eval=FALSE}
SSplotRunstest(base.model01,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model1,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model2,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model3,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model4,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model5,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model6,
               subplots = "cpue",
               add=T)
SSplotRunstest(base.model7,
               subplots = "cpue",
               add=T)
```

@@

```{r}
hci = SSplotHCxval(retroSummary1, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```

```{r eval=FALSE}
hci = SSplotHCxval(retroSummary01, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary1, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary2, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary3, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7) 
hci = SSplotHCxval(retroSummary4, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary5, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary6, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
hci = SSplotHCxval(retroSummary7, 
                   add = T, 
                   verbose = F, 
                   legendcex = 0.7)
```

@@

```{r}
mvln = SSdeltaMVLN(base.model1,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

```{r eval=FALSE}
mvln = SSdeltaMVLN(base.model01,
                   Fref = "Btgt", 
                   plot = TRUE,
                   addprj=TRUE,
                   virgin = TRUE,
                   mc=100)

@@

## Likelihood

@@

@@

# addWorksheet(wb, "variable_s2")
# writeData(wb, "variable_s2", outps1)
# 
# # Guardar el workbook
# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
# 
# 
# out_s2 <- read_excel("DataKrill.xlsx", 
#     sheet = "variable_s2")
outps1 %>%
  kbl(booktabs = TRUE,
      format = "html",
    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
  kable_paper("hover", 
              full_width = TRUE)%>%
  kable_styling(latex_options = c("striped",
                                  "condensed"),
                full_width = FALSE,
                font_size=9)#%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

@@

SSplotComparisons(summaryoutput01_1_7,
                  legendlabels = c(
                                   "No Spatial",

@@

```{r fig.height=6, fig.width=8}

@@

```{r fig.width=8, fig.height=7}
recs01 <- base.model01$recruit$pred_recr
recs1 <- base.model1$recruit$pred_recr
recs2 <- base.model2$recruit$pred_recr
recs3 <- base.model3$recruit$pred_recr
recs4 <- base.model4$recruit$pred_recr
recs5 <- base.model5$recruit$pred_recr
recs6 <- base.model6$recruit$pred_recr
recs7 <- base.model7$recruit$pred_recr

@@

```{r fig.width=8, fig.height=7}
dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S01")
dev1 <- base.model1$recruit[1:43,c(1,7)]%>% 
  mutate(Serie = "S1")
dev2 <- base.model2$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S2")
dev3 <- base.model3$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S3")
dev4 <- base.model4$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S4")
dev5 <- base.model5$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S5")
dev6 <- base.model6$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S6")
dev7 <- base.model7$recruit[1:43,c(1,7)] %>% 
  mutate(Serie = "S7")
